\documentclass[a4paper, 11pt]{article}

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{libertine} 

%% Sets page size and margins
\usepackage[a4paper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

%% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{booktabs} 
\usepackage{multirow}

\title{Problem Set 3}
\author{Huy Le Quang}

\begin{document}
\maketitle

\section {A long run perspective on the economy}

\textbf{Question 1:} What determines the amount of output an economy produces?

\textbf{Answer:} 

The amount of output an economy can produce depends on the two most important input factors: Labor and Capital. In addition, another determinant is the production technology that determines how much firms can produce given amounts of capital (K) and labor (L).

Production function: $Y = F(K,L)$ \\

\textbf{Question 2:} Explain how a competitive, profit-maximizing firm decides how much of each factor of production to demand.

\textbf{Answer:} 

Profit = Revenue - Costs

$\pi = P * F(K,L) - wL - iK$

FOC:

$P * \frac{dF}{dK} = P * MPK = i \rightarrow MPK = r$
$P * \frac{dF}{dL} = P * MPL = w \rightarrow MPL = \frac{w}{P}$

The firms will demand the quantity of labor which equates the Marginal Product of Labor (MPL) and real wage, similarly, they will demand the quantity of capital which equates the Marginal Product of Capital (MPK) and rent. \\

\textbf{Question 3:} Consider an economy with the following Cobb–Douglas production function:
$Y = AK^{1/3}L^{2/3}$; $A=1$.
Derive the equations describing capital demand (= demand for loanable funds = investment function) in this economy as a function of the real interest rate and labour.

\textbf{Answer:}

Profit maximizing firm considers:
$\pi = P * F(K,L) - wL - iK$

FOC:

$\frac{1}{3}K^{-2/3}L^{2/3} = r$
$\rightarrow K^{-2/3} = 3rL^{-2/3}$
$\rightarrow K = 3rL$ \\


\textbf{Question 4:} Using the above version of the Cobb–Douglas production function, calculate the capital income and the labour income in the economy. Interpret your results.

\textbf{Answer:}

Capital income = MPK * K = $\frac{1}{3}K^{1/3}L^{2/3} = \frac{1}{3}Y$

Labor Income = MPL * L = $\frac{2}{3}K^{1/3}L^{2/3} = \frac{2}{3}Y$

Capital accounts for 1/3 income and Labor accounts for 2/3 income. \\


\textbf{Question 5:} What determines consumption and investment?

\textbf{Answer:}
Real interest rate \\

\textbf{Question 6:}: Explain the difference between government purchases and transfer payments. Give two examples of each.

\textbf{Answer:} 

Government purchases are a measure of the dollar value of goods and services purchased directly by the government. For example, the government buys missiles and tanks, builds roads, and provides services such as air traffic control. All of these activities are part of GDP. Transfer payments are government payments to individuals that are not in exchange for goods or services. They are the opposite of taxes: taxes reduce household disposable income, whereas transfer payments increase it. Examples of transfer payments include Social Security payments to the elderly, unemployment insurance, and veterans' benefits.
\\

\textbf{Question 7:} What makes the demand for the economy’s output of goods and services equal the supply?

\textbf{Answer:} 
The real interest rate is the only endogenous variable in the model it must adjust to ensure that the demand for goods equals the supply. \\

\textbf{Question 8:} Explain what happens to consumption, investment, and the interest rate when the government increases taxes.

\textbf{Answer:} 

When the Government increases tax, and Y = const, the disposable income $\downarrow$

$\rightarrow$ Consumption decrease: $\delta C(\overline{Y} - \overline{T}) = \delta T * (-MPC)$

$\rightarrow$ National saving increase: S = Y - C - G $\rightarrow$ Supply for loanable fund shifts to the right

$\rightarrow$ Interest rate $\downarrow$ and Investment $\uparrow$
\\

\textbf{Question 9:} The government raises taxes by \$100 billion. If the marginal propensity to consume is 0.6, what happens to the following? Do they rise or fall? By what amounts?
a) Public saving.
b) Private saving.
c) National saving.
d) Investment.

\textbf{Answer:} 
a. Public saving increases by \$100 billion

b. Private saving increase by 100bil * 0.6 = 60bil

c. National saving increase by 160bil

d. Investment: increase, however, we cannot say the exact amount because it depends on the elasticity of the demand curve for loanable fund. \\


\textbf{Question 10:} Suppose that an increase in consumer confidence raises consumers’ expectations of future income and thus the amount they want to consume today. This might be interpreted as an upward shift in the consumption function. How does this shift affect investment and the interest rate?

\textbf{Answer:} 

Consumption increase $\rightarrow$ Saving decreases $\rightarrow$ Supply for loanable fund shifts to the left $\rightarrow$ interest rate increase and Investment decrease
 \\

\textbf{Question 11:} Consider an economy described by the following equations:
$$Y=C+I+G$$
$$Y=5,000$$
$$G=1,000$$
$$T=1,000$$
$$C=250 +0.75(Y −T)$$
$$I=1,000 −50r$$

\textbf{Answer:} 

a) In this economy, compute private saving, public saving, and national saving. \\
$S^{P} = Y - C - T = 5000 - 250 - 0.75(5000 - 1000) - 1000 = 750$

$S^{N} = T - G = 1000 - 1000 = 0$

$S = S^{P} + S^{N} = 750$

b) Find the equilibrium interest rate. \\
$Y = C + I + G$
$5000 = 3250 + 1000 - 50r + 1000$
$\rightarrow r = 5$

c) Now suppose that G rises to 1,250. Compute private saving, public saving, and national saving.\\
$S^{P} = 750$
$S^{N} = -250$
$S = 500$

d) Find the new equilibrium interest rate.\\
$Y = C + I + G$
$5000 = 3250 + 1000 - 50r + 1250$
$\rightarrow r = 10$ 
\\

\textbf{Question 12:} Suppose that the government increases taxes and government purchases by equal amounts. What happens to the interest rate and investment in response to this balanced-budget change? Does your answer depend on the marginal propensity to consume?

\textbf{Answer:} 

If the Government increase taxes and government purchases by the same amount, $S^{N}$ will stay the same. However, increasing taxes will lower consumption. And suppose that the government institutes an investment tax credit only for business investment. As $MPC < 1$, $S^{P}$ increases. Total saving S will increase, the Supply for loanable funds shifts to the right, lowering interest rates. MPC is an important variable to determine by how much $S^{P}$ will increase. The higher the MPC, the lower the increase of $S^{P}$. \\

\textbf{Question 13:} When the government subsidizes investment, such as with an investment tax credit, the subsidy often applies to only some types of investment. This question asks you to consider the effect of such a change. Suppose there are two types of investment in the economy: business investment and residential investment.

\textbf{Answer:}

Investment tax credit: An amount that businesses are allowed by law to deduct from their taxes, reflecting an amount they reinvest in themselves.

a) How does this policy affect the demand curve for business investment and the demand curve for residential investment?

The demand curve for business investment will shift to the right due to higher demand since it is now relatively cheaper to invest.

The demand curve for residential investment stays the same. \\

b) Draw the economy’s supply and demand for loanable funds. How does this policy affect the supply and demand for loanable funds? What happens to the equilibrium interest rate?
The supply for loanable funds stays the same, the demand for loanable funds shift to the right, increasing interest rates \\

c) Compare the old and the new equilibrium. 
As the supply for loanable fund stays the same, only demand for loanable fund shifts right, $\rightarrow$ quantity of investment stays the same, but interest rate is higher. \\

d) How does this policy affect the total quantity of investment? The quantity of business investment? The quantity of residential investment?

This policy does not change the quantity of investment because it does not change the supply for loanable fund. The gain from investment tax credit will be offset by the lose from higher interest rate. \\

\textbf{Question 14:} If consumption depended on the interest rate, how would that affect the conclusions reached in this chapter about the effects of fiscal policy?

\textbf{Answer:}  \\

In this chapter, we concluded that an increase in government expenditures reduces
national saving and raises the interest rate; it therefore crowds out investment by the full
amount of the increase in government expenditure. Similarly, a tax cut increases
disposable income and hence consumption; this increase in consumption translates into a
fall in national saving—again, it crowds out investment by the full amount of the increase
in consumption.

If consumption depends on the interest rate, then these conclusions about fiscal policy
are modified somewhat. If consumption depends on the interest rate, then so does saving.
The higher the interest rate, the greater the return to saving. Hence, it seems reasonable
to think that an increase in the interest rate might increase saving and reduce
consumption. \\

\section{Saving and Investment in a Small Open Economy}

\textbf{Question 15:} If a small open economy cuts defence spending, what happens to saving, investment, the trade balance, and the interest rate?

\textbf{Answer:}  \\

If a small open economy cuts defence spending, leaving other variables unchanged, $S^{G}$ increases, and national savings will increase. 
However, the interest rate does not change, because it is fixed at the world interest rate level, Investment stays the same as before.
The end result is that $S^{N} - I > 0$ $\rightarrow NX > 0$ $\rightarrow$ Net capital outflows > 0 (net lender to the world). \\

\textbf{Question 16:} If a small open economy bans the import of Japanese DVD players, what happens to saving, investment, the trade balance, and the interest rate?

\textbf{Answer:} \\
If a small open economy bans the import of Japanese DVD player, leaving other variables unchanged, NX increases, while interest rate $r^{*}$ does not change, Investment does not change. As $NX = S^{N} - I$ increases, and $\delta I = 0$, $S^{N}$ will increase. \\

\textbf{Question 17:} What would happen to the trade balance of a small open economy when government purchases increase, such as during a war? Does you answer depend on whether it is a local war or a world war?

\textbf{Answer:} \\
When Government purchases increase, the national saving will decrease. If it is a local war, the world interest rate does not change, Investment will stay the same, and Net export will decrease.

If it is a world war, the interest rate likely changes. If interest rate increases, Investment will decrease. If the decrease in national saving equals the decrease in investment, there will be no change in net export. However, as we do not know the magnitude of the interest rate change and the direction of change, we cannot conclude anything in this case. \\

\textbf{Question 18:} Suppose that some foreign countries begin to subsidize investment by instituting investment tax relief.

\textbf{Answer:} \\
If some foreign countries begin to subsidize investment by instituting investment tax credit, 

a) What happens to world investment demand as a function of the world interest rate? Investment demand will increase

b) What happens to the world interest rate? If investment demand increase, the world interest rate will also increase.

c) What happens to investment in our small open economy? As the world interest rate increases, the investment in the small open economy will decrease.

d) What happens to our trade balance? If other variables in the economy stays the same and investment decrease $\rightarrow S^{N} - I$ will increase and Net export will increase. \\

\section{The Determinants of the Real Exchange Rate}

\textbf{Question 19:} Use the model of the small open economy to predict what would happen to the trade balance, the real exchange rate, and the nominal exchange rate in response to each of the following events.

\textbf{Answer:} \\
a) A fall in consumer confidence about the future induces consumers to spend less and save more.

If consumers decide to spend less and save more, Saving will increase, and Net export will increase, the real exchange rate decrease. As NER depends both on RER and price level between two countries, and we do not have information about the change in price levels of two countries, we cannot conclude anything about the NER.
 
b) A tax reform increases the incentive for businesses to build new factories.

Investment increase, and Net export decreases. This leads to a decrease in RER.

c) The introduction of a stylish line of Toyotas makes some consumers prefer foreign cars over domestic cars.

Net export decreases, RER decreases.

d) The central bank doubles the money supply.

Inflation increases, Price level of the domestic country increases so RER increase. As RER increases, the price of domestic products are relatively more expensive than price of foreign products, NX will decrease.

e) New regulations restricting the use of credit cards increase the demand for money.

No change. \\

\textbf{Question 20:} Consider an economy described by the following equations:
$$Y = C + I + G + NX$$
$$Y = 5,000$$
$$G = 1,000$$
$$T = 1,000$$
$$C = 250 + 0.75(Y − T)$$
$$I = 1,000 − 50r$$
$$NX = 500 – 500\epsilon$$
$$r = r^{∗} = 5$$

\textbf{Answer:} 

a) In this economy, solve for national saving, investment, the trade balance, and the equilibrium exchange rate.
$$S^{N} = S^{P} + S^{G}$$
$$S^{N} = (Y-C-T) + (T-G)$$
$$S^{N} = 5000 - [250 + 0.75(5000 - 1000)] - 1000 + (1000 - 1000)$$
$$S^{N} = 750$$
$$I = 1000 - 50*5 = 750$$
$$NX = S^{N} - I = 500 - 500\epsilon = 0$$
$$\rightarrow \epsilon = 1$$
b) Suppose now that G rises to 1,250. Solve for national saving, investment, the trade balance, and the equilibrium exchange rate. Explain what you find.

Similar to a, we find that: 
$$S^{N} = 500$$
$$S^{N} - I = NX = -250$$
$$\rightarrow \epsilon = 1.5$$

National saving decreases, (S-I) curve shifts to the left, $\rightarrow$ NX decreases, and RER increases.

c) Now suppose that the world interest rate rises from 5 to 10 percent. (G is again 1,000.) Solve for national saving, investment, the trade balance, and the equilibrium exchange rate. Explain what you find.
$$S^{N} =750$$
$$I = 500$$
$$NX = 250$$
$$\rightarrow \epsilon = 0.5$$

I decreases, (S-I) curve shifts to the right, $\rightarrow$ NX increases, and RER decreases. \\


\textbf{Question 21:} What happens to the real exchange rate in exercises 15-18?

\textbf{Answer:} 

Q15: RER decreases

Q16: RER decreases

Q17: RER increases

Q18:  RER decreases


\textbf{Question 22:} In 2005, Federal Reserve Governor Ben Bernanke said in a speech: “Over the past decade a combination of diverse forces has created a significant increase in the global supply of saving—a global saving glut—which helps to explain both the increase in the U.S. current account deficit [a broad measure of the trade deficit] and the relatively low level of long-term real interest rates in the world today.” Is this statement consistent with the models you have learned? Explain.

\textbf{Answer:} 

Global saving increases $\rightarrow$ i/r decreases $\rightarrow$ Investment demand in US increase $\rightarrow$ (S-I) curve shift left $\rightarrow$ NX decreases (increase in US CA deficit). \\

\section{The Supply and Demand for Money}

\textbf{Question 23:} What does the assumption of constant velocity imply?

\textbf{Answer:} 

If we assume that the velocity of money is constant ($\bar{v}$), then the quantity equation becomes a theory of the effects of money supply:

$$M\bar{v} = PY$$ \\


\textbf{Question 24:} What is an inflation tax and who pays it?

\textbf{Answer:} 

Printing money to raise revenue is like imposing an inflation tax. Inflation causes a situation in which the money that the public holds in hand is less valuable as before. Different agents of the economy will be hurt (creditor, pensioners...) and others will gain (debtor). \\


\textbf{Question 25:} Name two channels which determine the demand for money. Explain.

\textbf{Answer:} 

Using Fisher equation and Quantity theory of money, we have:
$$(\frac{M}{P})^{D} = k[r + E(\pi)]Y$$

So the money demand depends on k (inverse of velocity: how much money people would like to hold), nominal interest rate and income. \\


\centering
END

\end{document}
